working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-03-12 Manuscript_draft/Manuscript_draft.Rmd
edits

For the understanding at what level or intensity the interference of pollen can occur is important to consider the relatedness of the interacting species. Closely related species are more likely to have similar traits @letten2015. The similarity in traits between closely related species can lead to higher chances of ovule usurpation/abortion @arceo2011 and therefore studies predict and show a greater negative effect of closely related species @ashman2013; @arceo2016. Few studies however, have focused on the impacts of heterospecific pollen on fitness of distantly related species @galen1989; @neiland1999. Despite the fact that far related species are also able to decrease species fitness (REFS). Yet, most insects and most stigmas have been found to carry multiple species of foreign pollen with little attention to degree of relatedness (Arceo-Gómez and Ashman (2016); Fang and Huang (2013). Understanding the role of foreign pollen from distantly related species thus deserves greater attention. The relatedness of foreign pollen gives a first snapshoot of where the pollen competition can occur and therefore could be a proxy of effect. With that said, @arceo2016 is the only work which has proven a greater effect of close related species through a meta-analysis but with low sample sizes and lack of significance. Therefore, there is a need to study the effect of heterospecific pollen of far and close related species at community level beyond single pairwise interactions.

Heterospecific pollen studies in nature have the complexity added of great environmental variability which can lead to confounding interpretations. Moreover, the great diversity of floral structures and small sizes of some of them such as Asteraceae species make tedious or almost impossible the studies of heterospecific pollen on the field. For this reason, we investigated how floral reproductive traits and relatedness mediate the impact of heterospecific pollen by creating an artificial co-flowering community in a glasshouse with 10 species belonging to three different families with different traits. Our study tries to answer the following questions: (i) Do heterospecific pollen reduce seed set? To what extent do (ii) floral reproductive traits and (iii) relatedness, mediate the impacts of heterospecific pollen on seed set.

@@

The study was conducted in a glasshouse at University of New England (Armidale, Australia) from November 2017 to March 2018. Rooms were temperature controlled depending on the requirements of the species with day and night temperature differences. The experimental design had species from three different families: Solanaceae, Brassicaceae and Convolvulaceae (**Table 1**). The species of the study had different reproductive traits and different degree of relatedness (**Figure 1**) where the reciprocal crosses between species allowed us to have multiple different scenarios of both traits and relatedness. Moreover, the species selected had fast life cycle and low structural flower complexity in order to perform the pollination treatments and grow the different species from seeds. For the purpose of the experiment all the species where considered as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14).

@@

**Table 1**
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
library(kableExtra)
Family <- c(rep("Brassicaceae",4), rep("Convolvulaceae", 2), rep("Solanaceae",4))

@@

The traits measured for each species were pollen per anther, number of ovules and stigma, style, ovary width and length. For the stigma, the stigmatic area was also measured and moreover the stigmas were divided in wet/dry type. All the morphometrical measurements were performed with a stereophotomicroscope. Pollen was counted for 20 anthers of each species with 4 replicates per sample with an hemocytometer. Previously, anthers were squashed on a known solution with the pippete tip and homogeneize with a vortex for 30 seconds. Ovule number was counted with the help of a stereomicroscope and a small grid over a petri dish from 15 randomly selected flowers. Fruits per number of flowers treated were counted for just Solanaceae species, we counted for all the species the number of seeds produced per average number of ovules. Levels of self-incompatibility were estimated by dividing the fruit set of hand self-pollination by hand cross pollination @lloyd1992.

>>>>

For the understanding at what level or intensity the interference of pollen can occur is important to consider the relatedness of the interacting species. Closely related species are more likely to have similar traits @letten2015. The similarity in traits between closely related species can lead to higher chances of ovule usurpation/abortion @arceo2011 and therefore studies predict and show a greater negative effect of closely related species @ashman2013; @arceo2016. Few studies however, have focused on the impacts of heterospecific pollen on fitness of distantly related species @galen1989; @neiland1999. Despite the fact that far related species are also able to decrease species fitness (REFS). Yet, most insects and most stigmas have been found to carry multiple species of foreign pollen with little attention to degree of relatedness (Arceo-Gómez and Ashman (2016); Fang and Huang (2013). Understanding the role of foreign pollen from distantly related species thus deserves greater attention. The relatedness of foreign pollen gives a first snapshoot of where the pollen competition can occur and therefore could be a proxy of effect. With that said, @arceo2016 is the only work until our knowledge which has proven a greater effect of close related species through a meta-analysis but with low sample sizes and lack of significance. Therefore, there is a need to study the effect of heterospecific pollen of far and close related species at community level beyond single pairwise interactions.

Heterospecific pollen studies in nature have the complexity added of great environmental variability which can lead to confounding interpretations. Moreover, the great diversity of floral structures and small sizes of some of them such as Asteraceae species make tedious or almost impossible the studies of heterospecific pollen on the field. For this reason, we investigated how floral reproductive traits and relatedness mediate the impact of heterospecific pollen by creating an artificial co-flowering community in a glasshouse with 10 species belonging to three different families with heterogeneous reproductive traits. Our study tries to answer the following questions: (i) Do heterospecific pollen reduce seed set? To what extent do (ii) floral reproductive traits and (iii) relatedness, mediate the impacts of heterospecific pollen on seed set.

@@

The study was conducted in a glasshouse at University of New England (Armidale, Australia) from November 2017 to March 2018. Rooms were temperature controlled depending on the requirements of the species with day and night temperature differences. The experimental design had species from three different families: Brassicaceae, Convolvulaceae and Solanaceae (**Table 1**). The species of the study had different reproductive traits and different degree of relatedness (see phylogenetic tree, **Figure 1**) where the reciprocal crosses between species allowed us to have multiple different scenarios of both traits and relatedness. Moreover, the species selected had fast life cycle and low structural flower complexity in order to perform the pollination treatments and grow the different species from seeds. For the purpose of the experiment all the species where considered as pollen recipient and as pollen donor (see interaction matrix, fig 2). Species were watered once or twice per day and fertilized weekly (NPK 23: 3.95: 14) and the rooms of the glasshouse were temperature controlled with temperature oscillations between day and night.

@@

**Table 1** Species list with family and genus.
```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
library(kableExtra)
Family <- c(rep("Brassicaceae",4), rep("Convolvulaceae", 2), rep("Solanaceae",4))

@@

The traits measured for each species were pollen per anther, pollen size, number of ovules and stigma, style, ovary width and length. For the stigma, the stigmatic area was also measured and moreover the stigmas were divided in wet/dry type with the help of the stereomicroscope. All the morphometrical measurements were performed with a stereophotomicroscope with the exception of pollen size that was carried out with a light microscope. Pollen was counted for 20 anthers of each species with 4 replicates per sample with an hemocytometer. Previously, anthers were squashed on a known solution with the pippete tip and homogeneize with a vortex for 30 seconds. Ovule number was counted with the help of a stereomicroscope and a small grid over a petri dish from 15 randomly selected flowers. Fruits per number of flowers treated were counted for just Solanaceae species with fleshy fruit. For all the species we counted the number of seeds produced per average number of ovules. Levels of self-incompatibility were estimated by dividing the fruit set of hand self-pollination by hand cross pollination @lloyd1992.

### 2019-03-12 Manuscript_draft/Manuscript_draft.tex
edits

For the understanding at what level or intensity the interference of
pollen can occur is important to consider the relatedness of the
interacting species. Closely related species are more likely to have
similar traits Letten and Cornwell (2015). The similarity in traits
between closely related species can lead to higher chances of ovule
usurpation/abortion Arceo-Gómez and Ashman (2011) and therefore studies
predict and show a greater negative effect of closely related species
Ashman and Arceo-Gómez (2013); Arceo-Gómez and Ashman (2016). Few
studies however, have focused on the impacts of heterospecific pollen on
fitness of distantly related species Galen and Gregory (1989); Neiland
and Wilcock (1999). Despite the fact that far related species are also
able to decrease species fitness (REFS). Yet, most insects and most
stigmas have been found to carry multiple species of foreign pollen with
little attention to degree of relatedness (Arceo-Gómez and Ashman
(2016); Fang and Huang (2013). Understanding the role of foreign pollen
from distantly related species thus deserves greater attention. The
relatedness of foreign pollen gives a first snapshoot of where the
pollen competition can occur and therefore could be a proxy of effect.
With that said, Arceo-Gómez and Ashman (2016) is the only work which has
proven a greater effect of close related species through a meta-analysis
but with low sample sizes and lack of significance. Therefore, there is
a need to study the effect of heterospecific pollen of far and close
related species at community level beyond single pairwise interactions.

Heterospecific pollen studies in nature have the complexity added of
great environmental variability which can lead to confounding
interpretations. Moreover, the great diversity of floral structures and
small sizes of some of them such as Asteraceae species make tedious or
almost impossible the studies of heterospecific pollen on the field. For
this reason, we investigated how floral reproductive traits and
relatedness mediate the impact of heterospecific pollen by creating an
artificial co-flowering community in a glasshouse with 10 species
belonging to three different families with different traits. Our study
tries to answer the following questions: (i) Do heterospecific pollen
reduce seed set? To what extent do (ii) floral reproductive traits and
(iii) relatedness, mediate the impacts of heterospecific pollen on seed
set.

@@

The study was conducted in a glasshouse at University of New England
(Armidale, Australia) from November 2017 to March 2018. Rooms were
temperature controlled depending on the requirements of the species with
day and night temperature differences. The experimental design had
species from three different families: Solanaceae, Brassicaceae and
Convolvulaceae (\textbf{Table 1}). The species of the study had
different reproductive traits and different degree of relatedness
(\textbf{Figure 1}) where the reciprocal crosses between species allowed
us to have multiple different scenarios of both traits and relatedness.
Moreover, the species selected had fast life cycle and low structural
flower complexity in order to perform the pollination treatments and
grow the different species from seeds. For the purpose of the experiment
all the species where considered as pollen recipient and as pollen donor
(see interaction matrix, fig 2). Species were watered once or twice per
day and fertilized weekly (NPK 23: 3.95: 14).

@@

\textbf{Table 1}

@@

The traits measured for each species were pollen per anther, number of
ovules and stigma, style, ovary width and length. For the stigma, the
stigmatic area was also measured and moreover the stigmas were divided
in wet/dry type. All the morphometrical measurements were performed with
a stereophotomicroscope. Pollen was counted for 20 anthers of each
species with 4 replicates per sample with an hemocytometer. Previously,
anthers were squashed on a known solution with the pippete tip and
homogeneize with a vortex for 30 seconds. Ovule number was counted with
the help of a stereomicroscope and a small grid over a petri dish from
15 randomly selected flowers. Fruits per number of flowers treated were
counted for just Solanaceae species, we counted for all the species the
number of seeds produced per average number of ovules. Levels of
self-incompatibility were estimated by dividing the fruit set of hand
self-pollination by hand cross pollination Lloyd and Schoen (1992).

@@

Although we have found a positive correlation between phylogenetic
distance and heterospecific pollen effect, this results have to be
treated carefully. From our results we want to highlight that also far
related species can affect negatively fitness but the effect from close
related is species can also have important detrimental effects.
Moreover, the effect of close related species can be masked by the
possibility of hibridization as it occure between two of our species of
Brassica. Moreover, although different effect between distinct donors
can occur we want to note the importance on the traits of the recipient
that determine an homogeneous effect between donors. These different
traits will define how it will be the effect across the different
species independently of the nature of the donor. Although traditionally
the nature of the pollen donor have been very studied, the
recipient\ldots{}

@@

Ideas about pollen size in heterospecific pollen effect. (still have to
develop it more\ldots{})

Let's classify pollen size in three groups in order to understand the
interaction between pollen donor and recipient: 1) Donor pollen size
\textless{} Recipient pollen size 2) Donor pollen size = Recipient
pollen size 3) Donor pollen size \textgreater{} Recipient pollen size

Now I try to develop each part

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Donor pollen size \textless{} Recipient pollen size
\end{enumerate}

Effect:

\begin{itemize}
\item
  Donor's pollen could clogg the stigma
\item
  Chemical inhibition
\end{itemize}

Traits associated with bigger pollen of the recipient:

\begin{itemize}
\item
  Recipient's pollen have faster pollen tube growth (example with my
  data)
\item
  Reduction in number of ovules (Also with my species)
\item
  Big differences in pollen size can be traduced in low relatednes
  therefore less likely of pollen germination on a far related stigma.
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  Donor pollen size = Recipient pollen size
\end{enumerate}

\begin{itemize}
\item
  Very relatedness dependant this point
\item
  Similar probabilities of taken space on the stigma
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  Donor pollen size \textgreater{} Recipient pollen size
\end{enumerate}

Effect:

-In small stigmas big pollen grains can occupy great part of the
stigmatic area.

-small pollen grains can get embeded

IB: Think also on using tree analysis to test if hp effect depends on
complex trait combinations. Tree analysis are great when two different
strategies lead to the same outcome. This would never been pick up by
GLMs. The r package is party\{\} . You can see an example applied to
birds is Sol et al 2010 Science. Ask me if you want more details or code
examples.

>>>>

For the understanding at what level or intensity the interference of
pollen can occur is important to consider the relatedness of the
interacting species. Closely related species are more likely to have
similar traits Letten and Cornwell (2015). The similarity in traits
between closely related species can lead to higher chances of ovule
usurpation/abortion Arceo-Gómez and Ashman (2011) and therefore studies
predict and show a greater negative effect of closely related species
Ashman and Arceo-Gómez (2013); Arceo-Gómez and Ashman (2016). Few
studies however, have focused on the impacts of heterospecific pollen on
fitness of distantly related species Galen and Gregory (1989); Neiland
and Wilcock (1999). Despite the fact that far related species are also
able to decrease species fitness (REFS). Yet, most insects and most
stigmas have been found to carry multiple species of foreign pollen with
little attention to degree of relatedness (Arceo-Gómez and Ashman
(2016); Fang and Huang (2013). Understanding the role of foreign pollen
from distantly related species thus deserves greater attention. The
relatedness of foreign pollen gives a first snapshoot of where the
pollen competition can occur and therefore could be a proxy of effect.
With that said, Arceo-Gómez and Ashman (2016) is the only work until our
knowledge which has proven a greater effect of close related species
through a meta-analysis but with low sample sizes and lack of
significance. Therefore, there is a need to study the effect of
heterospecific pollen of far and close related species at community
level beyond single pairwise interactions.

Heterospecific pollen studies in nature have the complexity added of
great environmental variability which can lead to confounding
interpretations. Moreover, the great diversity of floral structures and
small sizes of some of them such as Asteraceae species make tedious or
almost impossible the studies of heterospecific pollen on the field. For
this reason, we investigated how floral reproductive traits and
relatedness mediate the impact of heterospecific pollen by creating an
artificial co-flowering community in a glasshouse with 10 species
belonging to three different families with heterogeneous reproductive
traits. Our study tries to answer the following questions: (i) Do
heterospecific pollen reduce seed set? To what extent do (ii) floral
reproductive traits and (iii) relatedness, mediate the impacts of
heterospecific pollen on seed set.

@@

The study was conducted in a glasshouse at University of New England
(Armidale, Australia) from November 2017 to March 2018. Rooms were
temperature controlled depending on the requirements of the species with
day and night temperature differences. The experimental design had
species from three different families: Brassicaceae, Convolvulaceae and
Solanaceae (\textbf{Table 1}). The species of the study had different
reproductive traits and different degree of relatedness (see
phylogenetic tree, \textbf{Figure 1}) where the reciprocal crosses
between species allowed us to have multiple different scenarios of both
traits and relatedness. Moreover, the species selected had fast life
cycle and low structural flower complexity in order to perform the
pollination treatments and grow the different species from seeds. For
the purpose of the experiment all the species where considered as pollen
recipient and as pollen donor (see interaction matrix, fig 2). Species
were watered once or twice per day and fertilized weekly (NPK 23: 3.95:
14) and the rooms of the glasshouse were temperature controlled with
temperature oscillations between day and night.

@@

\textbf{Table 1} Species list with family and genus.

@@

The traits measured for each species were pollen per anther, pollen
size, number of ovules and stigma, style, ovary width and length. For
the stigma, the stigmatic area was also measured and moreover the
stigmas were divided in wet/dry type with the help of the
stereomicroscope. All the morphometrical measurements were performed
with a stereophotomicroscope with the exception of pollen size that was
carried out with a light microscope. Pollen was counted for 20 anthers
of each species with 4 replicates per sample with an hemocytometer.
Previously, anthers were squashed on a known solution with the pippete
tip and homogeneize with a vortex for 30 seconds. Ovule number was
counted with the help of a stereomicroscope and a small grid over a
petri dish from 15 randomly selected flowers. Fruits per number of
flowers treated were counted for just Solanaceae species with fleshy
fruit. For all the species we counted the number of seeds produced per
average number of ovules. Levels of self-incompatibility were estimated
by dividing the fruit set of hand self-pollination by hand cross
pollination Lloyd and Schoen (1992).

@@

Open with the main results\ldots{}

Although we have found a positive correlation between phylogenetic
distance and heterospecific pollen effect, this results have to be
treated carefully. From our results we want to highlight that also far
related species can affect negatively fitness but the effect from close
related is species can also have important detrimental effects.
Moreover, the effect of close related species can be masked by the
possibility of hibridization as it occure between two of our species of
Brassica. Moreover, although different effect between distinct donors
can occur we want to note the importance on the traits of the recipient
that determine an homogeneous effect between donors as we have shown in
figure xxx. These different traits will define how it will be the effect
across the different species independently of the nature of the donor
and also differently even between species of . Although traditionally
the nature of the pollen donor have been very studied, the
recipient\ldots{}

@@
